### 2019-03-14 Bucketing/intro.tex
A bunch of bucketing metadata, and new intro

Functional programming claims to be more ``mathematically tractable'' than other
``conventional'' approaches, due to properties like \emph{referential
transparency} and \emph{immutability}, which allow equational reasoning without
many of the complications caused by side effects and strict evaluation
order~\cite{hughes1989functional}. Languages such as Haskell combine these
advantages with static type systems to restrict the possible values of every
expression, making programs amenable to powerful analysis tools such as property
checking~\cite{claessen2011quickcheck} and theorem
proving~\cite{rosen2012proving}.

Whilst such tools can support or refute user-provided properties, we are
interested in the related problem of \emph{discovering} novel properties
automatically; a task addressed by the field of \emph{Mathematical Theory
  Exploration}. Whilst clever algorithms are used to guide their search,
existing MTE tools, including those applied to Haskell, ultimately rely on
enumerating all (well-typed) expressions. This limits their usefulness, as it is
only feasible to explore a small collection of functions at a time, relying on
the user to select promising combinations. We investigate the use of machine
learning methods to perform this selection automatically, removing the need for
user intervention whilst maintaining the ability to discover desirable
properties.

@@

\begin{enumerate}
  \item The application of machine learning algorithms to theory exploration,
    for intelligently discovering interesting sub-sets of Haskell libraries,
    which are more tractable to explore.
  \item A novel feature extraction method for transforming Haskell expressions
    into a form amenable to off-the-shelf learning algorithms.
  \item An implementation of these feature extraction and theory exploration
    approaches.
  \item A comparison of our methods with existing approaches, both for theory
    exploration in Haskell, and for machine learning in other languages.
\end{enumerate}

We describe our treatment of Haskell expressions in \S~\ref{sec:haskell} and
describe the \qspec{} theory exploration system we build upon in
\S~\ref{sec:theoryexploration}. We discuss our contributions in more depth in
\S~\ref{sec:contributions}, and provide implementation details in
\S~\ref{sec:implementation}. A variety of related work is surveyed in
\S~\ref{sec:related}, we briefly evaluate our implementations in
\S~\ref{sec:evaluation} and give several potential directions for future
research in \S~\ref{sec:future}.

>>>>

\emph{Theory exploration} (TE) describes an open-ended search for
interesting properties (conjectured or proven) of a given set of formal
definitions, such as a mathematical theory, a scientific model or a software
library. It has been applied to generating ``background lemmas'' in the
interactive proof assistant \textsc{Isabelle/HOL}~TODO{CITE ISACOSY AND HIPSTER}
and for automated theorem proving~TODO{CITE PAPER WHICH SHOWS IMPROVED PROVING},
as well as for generating specifications for (pure functional) programming
libraries~TODO{CITE QUICKSPEC 1, 2 AND SPECULATE}.

Despite the clever search strategies employed by these tools, they all rely on
enumerating combinations of definitions in some way, which can cause infeasible
running times on larger inputs (in our experience, a few dozen definitions). To
avoid this, users of these systems must carefully select only small subsets of
their definitions to explore at a time.

We dub such selection the \emph{bucketing problem}, and analyse its effect on
the running time of theory exploration tools on a particular problem set, and on
the quality of the statements they are able to generate. An automated solution
to the bucketing problem would accept arbitrarily many definitions, and
efficiently select sub-sets which are small enough to reasonably explore whilst
still enabling interesting properties to be discovered.

@@

\begin{enumerate}
  \item Defining the \emph{bucketing problem} for theory exploration.
  \item Statistical analysis of theory exploration performance on a large
    problem set, derived from existing theorem proving problem sets.
  \item Empirical analysis of the effect of bucketing on the quantity of
    desirable properties attainable by theory exploration.
  \item Experimental application of unsupervised machine learning to the
    bucketing problem.
  % \item A novel feature extraction method for transforming Haskell expressions
  %   into a form amenable to off-the-shelf learning algorithms.
  % \item An implementation of these feature extraction and theory exploration
  %   approaches.
  % \item A comparison of our methods with existing approaches, both for theory
  %   exploration in Haskell, and for machine learning in other languages.
\end{enumerate}

We describe theory exploration in more detail, as well as the \qspec{} tool we
use in our experiments, in \S~\ref{sec:theoryexploration}. We introduce the
bucketing problem in \S~\ref{sec:bucketing} and measure its effect on theory
exploration systems in \S~\ref{sec:implementation}. We attempt (unsuccessfully)
to apply machine learning to the bucketing problem in \S~\ref{sec:clustering}.
A variety of related work is surveyed in \S~\ref{sec:related} and we conclude in
\S~\ref{sec:conclusion} with potential directions for future research.
